---
title: "hw11"
author: "Jiayi"
date: "2024-12-07"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(faraway)
```

# Q1 (a)
```{r}
# Data
soybean <- c(158, 171, 193, 199, 230, 243, 248, 248, 250, 267, 271, 316, 327, 329)
linseed <- c(141, 148, 169, 181, 203, 213, 229, 244, 257, 260, 271, 309)

z <- c(soybean, linseed)
group <- c(rep(1, length(soybean)), rep(2, length(linseed)))

# Kolmogorov-Smirnov statistic function
ks_statistic <- function(x, group) {
  Fn <- ecdf(x[group == 1]) # Empirical CDF for group 1
  Gm <- ecdf(x[group == 2]) # Empirical CDF for group 2
  max(abs(Fn(x) - Gm(x))) 
}

# Observed statistic
obs_stat <- ks_statistic(z, group)

# Randomization distribution
set.seed(123)  # Ensure reproducibility
perm_stats <- replicate(10000, {
  perm_group <- sample(group)          # Randomize group labels
  ks_statistic(z, perm_group)          # Compute KS statistic for permuted data
})

# Plot the randomization distribution
hist(perm_stats, main = "Randomization Distribution of KS Statistic",
     xlab = "KS Statistic", col = "lightblue")
abline(v = obs_stat, col = "red", lwd = 2, lty = 2) 
```

# Q1 (b)
```{r}
# Compute p-value
p_value <- mean(perm_stats >= obs_stat)
cat("The p-value is:", p_value)
```

# Q2 (a)
```{r}
# Load the data
data(alfalfa, package = "faraway")

# Fit the model
model <- aov(yield ~ inoculum + shade + irrigation, data = alfalfa)

# Summary of the model
summary(model)
```
From the ANOVA table:

The treatment factor *inoculum* has a significant effect on yield (F = 12.709, p = 0.000284), as the p-value is *less than the significance level of 0.05*.


The blocking factor *shade* also has a significant effect on yield (F = 7.125, p = 0.003533), as its p-value is *below 0.05*.


The blocking factor *irrigation*, however, does not have a significant effect on yield (F = 1.350, p = 0.307872), as its *p-value is greater than 0.05*.

Thus, the treatment factor and the shade blocking factor are significant, while the irrigation blocking factor is not.


# Q2 (b)
```{r}
# Tukey's Honest Significant Difference Test
tukey_results <- TukeyHSD(model, "inoculum", conf.level = 0.95)

# Display results
print(tukey_results)

# Significant pairwise differences
significant <- tukey_results$inoculum[tukey_results$inoculum[, "p adj"] < 0.05, ]
significant
```

From the Tukey HSD test results:


Pairwise comparisons reveal that *inoculum E* is significantly different from several other levels *at an experiment-wise significance level of 0.05*:


	•	E-A: p = 0.0005166
	
	•	E-B: p = 0.0014163
	
	•	E-C: p = 0.0005764
	
	•	E-D: p = 0.0017334
	
	
No other pairwise comparisons show significance as their adjusted p-values exceed 0.05. In conclusion, inoculum level E stands out as being significantly different from levels A, B, C, and D, while the other pairwise differences are not statistically significant.