---
title: "hw12"
author: "Jiayi"
date: "2024-12-14"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1.(a)
```{r}
# Load the necessary library
library(faraway)

# Load the dataset
data(denim)

fixed_model <- lm(waste ~ supplier, data = denim)
summary(fixed_model)

anova(fixed_model)

```

The p-value is 0.334, which is much larger than 0.05. Therefore, the supplier is not significant at 0.05.

## Q1.(b)
```{r}
library(lme4)

# Fit a mixed model with supplier as a random effect
random_model <- lmer(waste ~ 1 + (1 | supplier), data = denim)
summary(random_model)

# Estimated variance components
VarCorr(random_model)
```
The estimated variance components for the random effects model are as follows:

	•	Variance for supplier: 0.6711
	
	•	Residual variance: 97.3350

This indicates that the variability in waste is primarily due to residual variance, with a small contribution from variability between suppliers.

## Q1.(c)
```{r}
# Compute ICC
variance_components <- as.data.frame(VarCorr(random_model))
icc <- variance_components$vcov[1] / sum(variance_components$vcov)
icc
```
The intraclass correlation is 0.006847128.

## Q1.(d)
```{r}
library(boot)

# Parametric bootstrap for Supplier variance
set.seed(123)
boot_res <- bootMer(random_model, FUN = function(x) {
  as.numeric(VarCorr(x)$supplier[1, 1])
}, nsim = 1000, use.u = TRUE)

# Bootstrap p-value
observed_variance <- VarCorr(random_model)$supplier[1, 1]
p_value <- mean(boot_res$t >= observed_variance)
p_value
```
The p-value is 0.345.

## Q2.(a)
```{r}
# Load the dataset
data(eggprod)

# Fit the mixed model
egg_model <- lmer(eggs ~ treat + (1 | block), data = eggprod)
summary(egg_model)

library(emmeans)
best_treatment <- emmeans(egg_model, "treat")
print(best_treatment)
```
Treatment E is the best treatment, with significantly higher egg production compared to *Treatment O* (t=-3.056 gives a p-value 0.015 smaller than 0.05). Also, The confidence intervals for Treatments E and O do not overlap, which confirms the difference between them is statistically significant.


However, Treatment E is *not significantly better than Treatment F* (t=-0.449 gives a p-value 0.667  smaller than 0.05). Also, the confidence intervals for Treatments E and F overlap, which means the difference between them is not statistically significant.

## Q2.(b)
```{r}
# Test the significance of block effects using likelihood ratio test
null_model <- lm(eggs ~ treat, data = eggprod)
anova(egg_model, null_model)
```
The significant level is 0.4071, which is much larger than 0.05. This means the random effects for blocks are not significant.
