---
title: "hw2"
author: "Jiayi"
date: "2024-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r loadData}
# I download the cship database on my computer. Put it in the same folder with my code. Then, I read csv directly.
cship <- read.csv("cship.csv")
head(cship)
```

Q1:
```{r loadData}
# Construct the design matrix X using Age, Tonnage, and Length as explanatory variables with an intercept
X <- as.matrix(cbind(1, cship$Age, cship$Tonnage, cship$Length)) # cbind means column binding things together

head(X)
```

Q2:Compute X′X and X′y: two matrices are printed out
```{r X′X}
XtX <- t(X) %*% X
y <- cship$passengers
Xty <- t(X) %*% y

XtX
Xty
```

Q3: Find the eigenvalues of X′X
```{r eigenvalue}
eigenvalues <- eigen(XtX)$values
eigenvalues
```

Q4: Compute the inverse of X′X
```{r inverse}
XtX_inv <- solve(XtX)
XtX_inv

```
Q5: Find the column rank of  X
```{r column_rank}
rank_X <- qr(X)$rank
rank_X
```

Q6: Compute βˆ (the regression coefficients)
```{r regression_coefficient}
beta_hat <- XtX_inv %*% Xty
beta_hat
```
Q7: 
The first value in beta_hat is the intercept, and the remaining values are the coefficients for Age, Tonnage, and Length. Each coefficient indicates the change in the predicted number of passengers for a one-unit increase in the corresponding variable, holding the other variables constant.

To illustrate, for each additional year of age of the ship, the number of passengers is expected to increase by approximately 11.27 passengers, assuming that the tonnage and length are held constant. For each additional *thousand tons* of ship weight (tonnage), the number of passengers is expected to increase by approximately 24.45 passengers passengers, assuming that age and length are held constant. For each additional *hundred feet* in the length of the ship, the number of passengers is expected to increase by approximately 34.06 passengers, assuming that age and tonnage are held constant.

Q8:
```{r prediction}
# Compute the average of each explanatory variable
x_avg <- colMeans(X)

# Predict the number of passengers for the average ship
y_avg_predicted <- sum(x_avg * beta_hat)
y_avg_predicted
```

Q9: Compute y_hat and find RSS
```{r y_hat}
# Compute the predicted values
y_hat <- X %*% beta_hat

# Compute the residual sum of squares (RSS)
RSS <- sum((y - y_hat)^2)

# Display RSS
RSS
```
Therefore, RSS is 1484.756.

Q10: Compute (σ_hat)^2
```{r σ_hat_square}
n <- length(y)
p <- ncol(X) - 1  # Number of explanatory variables
sigma2_hat <- RSS / (n - p - 1)

# Display sigma^2
sigma2_hat
```
Therefore, sigma squared is 9.641271. 

Q11: Covariance matrix of β_hat
```{r y_hat_covariance_matrix}
# Estimate the covariance matrix of beta hat
cov_beta_hat <- sigma2_hat * XtX_inv

# Display the covariance matrix
cov_beta_hat
```
Q12: Compute R^2
```{r r_squared}
# Compute the total sum of squares (TSS)
TSS <- sum((y - mean(y))^2)

# Compute the R-squared
R_squared <- 1 - (RSS / TSS)

R_squared
```
Therefore, R − Squared is 0.899013.

Q13: Find the ship with the largest residual
```{r largest_residual}
# Compute the residuals
residuals <- y - y_hat

# Find the ship with the largest residual
largest_residual_ship <- cship$ShipName[which.max(abs(residuals))] # using the index
```
"Queen Mary 2" is the ship with the largest residual.