---
title: "hw7"
author: "Jiayi"
date: "2024-10-28"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(faraway)
library(leaps)  
```

### Part 1: Prostate Data Analysis

#### Load and Explore the Data

```{r}
data(prostate)
str(prostate)
summary(prostate)
```

#### (a) Backward Elimination with Alpha = 0.20

```{r}
model_full <- lm(lpsa ~ ., data = prostate)
model_backward <- step(model_full, direction = "backward", k = qchisq(0.20, 1, lower.tail = FALSE))
summary(model_backward)
```

#### (b), (c), (d) AIC, BIC, and Mallows Cp using leaps package

```{r}
leaps_result <- regsubsets(lpsa ~ ., data = prostate, nbest = 1)
summary_leaps <- summary(leaps_result)

# Number of observations
n <- nrow(prostate)

# Calculate AIC for each model
aic_values <- numeric(length(summary_leaps$rss))
for (i in 1:length(aic_values)) {
  k <- sum(summary_leaps$which[i, ])  # Number of predictors (including intercept)
  rss <- summary_leaps$rss[i]         # Residual Sum of Squares
  aic_values[i] <- n * log(rss / n) + 2 * k
}

# Get the selected models' indices
aic_model_index <- which.min(aic_values)
bic_model_index <- which.min(summary_leaps$bic)
cp_model_index <- which.min(summary_leaps$cp)


# Extract the variables included in those models
aic_model_vars <- names(which(summary_leaps$which[aic_model_index, ]))
bic_model_vars <- names(which(summary_leaps$which[bic_model_index, ]))
cp_model_vars <- names(which(summary_leaps$which[cp_model_index, ]))

# Report the selected variables
list(
  AIC_Model_Vars = aic_model_vars,
  BIC_Model_Vars = bic_model_vars,
  Cp_Model_Vars = cp_model_vars
)
```

#### (e) Stepwise Selection using AIC Criterion

```{r}
model_stepwise <- step(model_full, direction = "both", trace = 0)
summary(model_stepwise)
```
The stepwise selection method, which uses AIC to determine the best model, selected lcavol (log cancer volume), lweight (log prostate weight), age, lbph (log benign prostatic hyperplasia), and svi (seminal vesicle invasion) as the most relevant predictors for lpsa (log prostate-specific antigen)


### Part 2: TeenGamb Data Analysis

#### Load and Explore the Data

```{r}
data(teengamb)
str(teengamb)
summary(teengamb)
```


#### (a) Backward Elimination with Alpha = 0.20

```{r}
model_full_gamble <- lm(gamble ~ ., data = teengamb)
model_backward_gamble <- step(model_full_gamble, direction = "backward", k = qchisq(0.20, 1, lower.tail = FALSE))
summary(model_backward_gamble)
```

#### (b), (c), (d) AIC, BIC, and Mallows Cp using leaps package

```{r}
leaps_result_gamble <- regsubsets(gamble ~ ., data = teengamb, nbest = 1)
summary_leaps_gamble <- summary(leaps_result_gamble)

# Number of observations
n <- nrow(teengamb)

# Calculate AIC for each model
aic_values_gamble <- numeric(length(summary_leaps_gamble$rss))
for (i in 1:length(aic_values_gamble)) {
  k <- sum(summary_leaps_gamble$which[i, ])  # Number of predictors (including intercept)
  rss <- summary_leaps_gamble$rss[i]         # Residual Sum of Squares
  aic_values_gamble[i] <- n * log(rss / n) + 2 * k
}

# Get the model indices that minimize AIC, BIC, and Mallows Cp
aic_model_index_gamble <- which.min(aic_values_gamble)
bic_model_index_gamble <- which.min(summary_leaps_gamble$bic)
cp_model_index_gamble <- which.min(summary_leaps_gamble$cp)

# Extract the selected variables (removing intercept for clarity)
aic_model_vars_gamble <- names(which(summary_leaps_gamble$which[aic_model_index_gamble, ]))[-1]
bic_model_vars_gamble <- names(which(summary_leaps_gamble$which[bic_model_index_gamble, ]))[-1]
cp_model_vars_gamble <- names(which(summary_leaps_gamble$which[cp_model_index_gamble, ]))[-1]

# Print the selected variables for AIC, BIC, and Cp models
cat("AIC Model Variables:", paste(aic_model_vars_gamble, collapse = ", "), "\n")
cat("BIC Model Variables:", paste(bic_model_vars_gamble, collapse = ", "), "\n")
cat("Cp Model Variables:", paste(cp_model_vars_gamble, collapse = ", "), "\n")
```

#### (e) Stepwise Selection using AIC Criterion

```{r}
model_stepwise_gamble <- step(model_full_gamble, direction = "both", trace = 0)
summary(model_stepwise_gamble)
```
Three varibales: sex, income, verbal